---
title: "Sharday Study Habit Figures"
author: "Emily Driessen"
date: "6/1/2022"
output: html_document
---

This R Markdown file presents all of the codes used to make the figures for the quantitative analysis to answer the following research questions:

In this quantitative analysis, we explored the following research questions:

Research Questions:
1.	Are there differences in the use of study behaviors between lower-level and upper-level students?
2.	What is the effect of each study behaviors on student performance?
3.	How does exam score predict the study behaviors students use?
4.	How do student study behaviors change over the course of the semester?

We break down the analysis for each research question after conducting descriptive analyses. 

```{r}
#clear environment
rm(list=ls())
#set working directory
setwd("~/Box/Research Projects/Sharday's Study Habits")
#upload data
CombinedZ <- read.csv("~/Box/Research Projects/Sharday's Study Habits/CombinedZ.csv") #This dataset has Z-scored exam scores as normalized within each instructors' classes.

##Load nlme
library(nlme) #for lme4 use lmer for poisson distribution with random effects
library(tidyverse) #for pipes
library(lme4) 
library(dplyr)
library(tidyr)
library(ggplot2)
```

#Clean Data
```{r}
CombinedZ<-CombinedZ %>%
 mutate(instructoranon=as.character(instructoranon))%>%
mutate(instructoranon=fct_relevel(instructoranon, "A","B","C", "D")) %>% #mutate the instructor column by releveling it to have the desired order 
  filter(!is.na(Zexamscore), exam_score>0) %>% #removes any NA for Zexamscore, and removes any exam scores that were 0 before they were z-scored
  select(student_id, Zexamscore, exam_score, instructoranon, exam, gender, ethnicity, Level, active_strat_count, passive_strat_count, distracted_study, GPA_hs, enroll_year, class)%>%
rename(Student="student_id", Exam = "exam", Score = "exam_score", ZScore = "Zexamscore", Gender = "gender", Ethnicity = "ethnicity", Active = "active_strat_count", Passive = "passive_strat_count", Distracted = "distracted_study", GPA="GPA_hs", Enrolled="enroll_year", Class="class")
```


1.	Are there differences in the use of study behaviors between students in different courses?

```{r}
StrategyCountClass<-CombinedZ %>%
  select(Score, Student, instructoranon, Exam, Gender, Ethnicity, Class, Active, Passive)%>%
  pivot_longer(-c(Student, Score, Class, instructoranon, Exam, Gender, Ethnicity), "StudyHabitType") %>%
  rename("Count"="value")  #rename the default "value" column as "Percent" 


meanactive<-CombinedZ%>%
  group_by(Class) %>%
 summarise(mean(Active, na.rm = TRUE)) 
 # mean active strategies for lower = 2.79 and for upper is 3.13

meanpassive<-CombinedZ%>%
  group_by(Class) %>%
 summarise(mean(Passive, na.rm = TRUE)) 

meanpassive
 # mean passive strategies for A&P is 1.95, Genetics is 1.55, and microbiology is 1.89

#I can use this information to create trendlines for the plot. 

means <- StrategyCountClass %>%
       group_by(Class, StudyHabitType) %>%
       summarise(grp.mean=mean(Count)) 

means[is.na(means)] <- 1.95
       
  StratcountClass<-ggplot(StrategyCountClass, aes(x =Count, fill = Class)) +
    geom_bar(position = position_dodge(preserve = "single")) +
    scale_fill_manual(values = c("#91D1C2B2", "#F39B7FB2", "#FF7256" )) +
  scale_color_manual(values = c("#91D1C2B2", "#F39B7FB2", "#FF7256")) +
    scale_x_continuous(name = "# of study strategies used", breaks=c(0,1,2,3,4,5,6), limits=c(-0.5, 6.5)) + 
    scale_y_continuous(name = "# of students", breaks=c(0,25,50,75,100,125,150, 175, 200, 225, 250, 275, 300, 325, 350, 375, 400), limits=c(0, 400)) +  
    theme_bw() + 
   theme(
        panel.grid = element_blank())+ 
  facet_wrap(~StudyHabitType)+
    geom_vline(data=means, aes(xintercept=grp.mean, color=Class), linetype="dashed")
  
  StratcountClass

#library(tidyverse)

#https://stackoverflow.com/questions/43569732/how-to-add-legend-to-geom-vline-in-facet-histograms
 
#https://rsample.tidymodels.org/articles/Applications/Intervals.html

  
  #Now to Save!
# - vector based output height = 4.5, width = 6 inches
ggsave(StratcountClass, file=" Stratcount.png",
       height = 4.5,
       width = 6,
       units = "in",
       dpi = 300)

Active<-glmer(Active~Class+(1|instructoranon/Student), data=CombinedZ, family=poisson, na.action=na.exclude) #active study strategies is a count, so I used glmer to specify poisson as family

summary(Active)

confint(Active)

Passive<-glmer(Passive~Class+(1|instructoranon/Student), data=CombinedZ, family=poisson, na.action=na.exclude) #passive study strategies is a count, so I used glmer to specify poisson as family
summary(Passive)
confint(Passive)

##There are significant differences between courses for both active and passive. Specifically, Genetics is not different than anatomy, but microbiology is significantly different that anatomy when it comes to the number of active strategies used. For passive strategies, Microbiology is not different from anatomy, but genetics is different from anatomy. 
```

Are there differences in the use of study behaviors between students by instructor?
```{r}
StrategyCountInstructor<-CombinedZ %>%
  select(Score, Student, instructoranon, Exam, Gender, Ethnicity, Active, Passive)%>%
  pivot_longer(-c(Student, Score, instructoranon, Exam, Gender, Ethnicity), "StudyHabitType") %>%
  rename("Count"="value")  #rename the default "value" column as "Percent" 

meanactiveinstructor<-CombinedZ%>%
  group_by(instructoranon) %>%
 summarise(mean(Active, na.rm = TRUE)) 

meanactiveinstructor
 # mean active strategies for instructor A is 2.71, instructor B is 2.87, instructor C is 3.24, instructor D is 2.81

meanpassiveinstructor<-CombinedZ%>%
  group_by(instructoranon) %>%
 summarise(mean(Passive, na.rm = TRUE)) 

meanpassiveinstructor
 # mean passive strategies for instructor A is 2.13, instructor B is 1.77, instructor C is 1.89, instructor D is 1.55

#I can use this information to create trendlines for the plot. 

means <- StrategyCountInstructor %>%
       group_by(instructoranon, StudyHabitType) %>%
       summarise(grp.mean=mean(Count)) 

means[is.na(means)] <- 1.95
       
  StratcountInstructor<-ggplot(StrategyCountClass, aes(x =Count, fill = instructoranon)) +
    geom_bar(position = position_dodge(preserve = "single")) +
    scale_fill_manual(values = c("#91D1C2B2","#458B74", "#FF7256", "#F39B7FB2" )) +
  scale_color_manual(values = c("#91D1C2B2", "#458B74", "#FF7256", "#F39B7FB2")) +
    scale_x_continuous(name = "# of study strategies used", breaks=c(0,1,2,3,4,5,6), limits=c(-0.5, 6.5)) + 
    scale_y_continuous(name = "# of students", breaks=c(0,25,50,75,100,125,150, 175, 200, 225, 250, 275, 300, 325, 350, 375, 400), limits=c(0, 300)) +  
    theme_bw() + 
   theme(
        panel.grid = element_blank())+ 
  facet_wrap(~StudyHabitType)+
    geom_vline(data=means, aes(xintercept=grp.mean, color=instructoranon), linetype="dashed")
  
  StratcountInstructor

  
  #Now to Save!
# - vector based output height = 4.5, width = 6 inches
ggsave(StratcountInstructor, file=" StratcountInstructor.png",
       height = 4.5,
       width = 6,
       units = "in",
       dpi = 300)

Passive<-glmer(Passive~instructoranon+(1|Student), data=CombinedZ, family=poisson, na.action=na.exclude) #passive study strategies is a count, so I used glmer to specify poisson as family
summary(Passive)
confint(Passive)
##Should I keep student nested in instructor for this model when I'm testing the differences between instructors? The resuls of this analysis show that instructor B, C, and D are all statistically different from instructor A when it comes to the number of passive strategies students used. 

Active<-glmer(Active~instructoranon+(1|Student), data=CombinedZ, family=poisson, na.action=na.exclude) #active study strategies is a count, so I used glmer to specify poisson as family
summary(Active)
confint(Active)

##This analysis showed there is a difference between the number of active strategies used in instructor C's class and instructor A's class. However, instructor B and D are not different from Instructor A. 
```


2.What is the effect of each study behavior on student performance?
```{r}

CombinedZ1<-CombinedZ %>%
mutate(Activepercent = ((Active)/(Active + Passive)*100))%>% 
  select(Score, Student, instructoranon, Exam, ZScore, Level, Activepercent)

#Passive by class
ExamPassiveClass<-ggplot(CombinedZ, aes(x = Passive, y = Score)) + # map data
  geom_jitter(alpha=.3,) +
#scale_fill_manual(values = c("blue", "darkblue")) +
  #scale_color_manual(values = c("blue" ,"darkblue")) +
  scale_y_continuous(name = "Exam Score", breaks=c(20,30,40,50,60,70,80,90,100), limits=c(30, 100)) +
 scale_x_continuous(name = "Count passive study strategies", breaks=c(0,1, 2, 3, 4), limits=c(0, 4)) +   # CHANGE: change label
  theme_bw() + 
  theme(panel.grid = element_blank()) +
  facet_wrap(~Class) +
  geom_smooth(method=lm, se=FALSE, col='black', size=1)

ExamPassiveClass

ggsave(ExamPassiveClass, file="ExamPassiveClass.png",
       height = 4.5,
       width = 8,
       units = "in",
       dpi = 300)

#Passive
ExamPassive<-ggplot(CombinedZ, aes(x = Passive, y = Score)) + # map data
  geom_jitter(alpha=.3,) +
  scale_y_continuous(name = "Exam Score", breaks=c(20,30,40,50,60,70,80,90,100), limits=c(30, 100)) +
 scale_x_continuous(name = "Count passive study strategies", breaks=c(0,1, 2, 3, 4), limits=c(0, 4)) +   # CHANGE: change label
  theme_bw() + 
  theme(panel.grid = element_blank()) +
  geom_smooth(method=lm, se=FALSE, col='black', size=1)

ExamPassive

ggsave(ExamPassive, file="ExamPassive.png",
       height = 4.5,
       width = 8,
       units = "in",
       dpi = 300)


#Active by class
ExamActiveClass<-ggplot(CombinedZ, aes(x = Active, y = Score)) + # map data
  geom_jitter(alpha=.3,) +
  scale_y_continuous(name = "Exam Score", breaks=c(20,30,40,50,60,70,80,90,100), limits=c(30, 100)) +
 scale_x_continuous(name = "Count active study strategies", breaks=c(0,1, 2, 3, 4, 5, 6), limits=c(0, 6)) +   # CHANGE: change label
  theme_bw() + 
  theme(panel.grid = element_blank()) +
  facet_wrap(~Class) +
  geom_smooth(method=lm, se=FALSE, col='black', size=1)

ExamActiveClass

ggsave(ExamActiveClass, file="ExamActiveClass.png",
       height = 4.5,
       width = 8,
       units = "in",
       dpi = 300)

#Active
ExamActive<-ggplot(CombinedZ, aes(x = Active, y = Score)) + # map data
  geom_jitter(alpha=.3,) +
  scale_y_continuous(name = "Exam Score", breaks=c(20,30,40,50,60,70,80,90,100), limits=c(30, 100)) +
 scale_x_continuous(name = "Count active study strategies", breaks=c(0,1, 2, 3, 4, 5, 6), limits=c(0, 6)) +   # CHANGE: change label
  theme_bw() + 
  theme(panel.grid = element_blank()) +
  geom_smooth(method=lm, se=FALSE, col='black', size=1)

ExamActive

ggsave(ExamActive, file="ExamActive.png",
       height = 4.5,
       width = 8,
       units = "in",
       dpi = 300)

#Passive

##Passive by Class
PassivecountexamZ<-lme(ZScore~Passive + Class + Passive:Class, random=~1|instructoranon/Student, data=CombinedZ, na.action=na.exclude)

summary(PassivecountexamZ)

##Interaction is not significant, so I took that out and analyzed passive strategies together.
PassivecountexamZ<-lme(ZScore~Passive + Class, random=~1|instructoranon/Student, data=CombinedZ, na.action=na.exclude)

summary(PassivecountexamZ)

##Active by Class
ActivecountexamZ<-lme(ZScore~Active + Class + Active:Class, random=~1|instructoranon/Student, data=CombinedZ, na.action=na.exclude)

summary(ActivecountexamZ)

##No sig effect of the interaction, so analyzed active all together

ActivecountexamZ<-lme(ZScore~Active + Class, random=~1|instructoranon/Student, data=CombinedZ, na.action=na.exclude)

summary(ActivecountexamZ)


##There is a significant effect of active strategies on student performance. To make it more interperatable, I'm going to run these models with exam score in percent rather than z-scored. 

Passivecountexam<-lme(Score~Passive + Class, random=~1|instructoranon/Student, data=CombinedZ, na.action=na.exclude)

summary(Passivecountexam)
intervals(Passivecountexam)

Activecountexam<-lme(Score~Active + Class, random=~1|instructoranon/Student, data=CombinedZ, na.action=na.exclude)

summary(Activecountexam)
intervals(Activecountexam)

```

3. How do student study behaviors change over the course of the semester?

###Active by exam
```{r}
CountActiveExamnolevel<-CombinedZ %>%
  group_by(Exam) %>%
count(Active)%>% #This tells me the number of students who said each number of strategies for active by level. 
  rename("number_of_students"="n") %>% #rename the default "value" column as "Percent" 
  mutate(Total_of_students = number_of_students) %>%#I created a dummy column for total number of students. Then, I manually added up all of the lower students and all of the upper students. There are 968 exam 1 students and 900 exam 3 students. 
      mutate(Total_of_students=recode(Total_of_students,"45"="968","112"="968","192"="968", "237"="968", "217"="968", "111"="968","54"="968", "42"="900", "106"="900","197"="900","204"="900", "187"="900", "76"="900", "34"="900")) %>%
  mutate(Percent_of_students_Active = number_of_students) %>%#Create dummy column and fill in values
  mutate(Percent_of_students_Active=recode(Percent_of_students_Active,"45"="4.65","112"="11.57","192"="19.83", "237"="24.48", "217"="22.42", "111"="11.47","54"="5.58", "42"="4.67", "106"="11.78","197"="21.89","204"="22.67", "187"="20.78", "76"="8.44", "34"="3.78")) %>%
   mutate(Percent_of_students_Active=as.numeric(Percent_of_students_Active))%>%
 mutate(Total_of_students=as.numeric(Total_of_students))

meanactivenolevel<-CombinedZ%>%
  group_by(Exam) %>%
 summarise(mean(Active, na.rm = TRUE)) %>%
  rename(grp.mean="mean(Active, na.rm = TRUE)")

ActivePercentExamnolevel<-ggplot(CountActiveExamnolevel, aes(x=Active, y=Percent_of_students_Active, fill=Exam)) +
    geom_col(position = position_dodge(preserve = "single")) +
        scale_fill_manual(values=c("#8491B4B2", "#7E6148B2")) +
   scale_color_manual(values = c("#8491B4B2", "#7E6148B2")) +
    scale_x_continuous(name = "# of study strategies used", breaks=c(0,1,2,3,4,5,6), limits=c(-0.5, 6.5)) + 
    scale_y_continuous(name = "% of students", breaks=c(0, 10, 20, 30, 40), limits=c(0, 45)) +  
    theme_bw() + 
   theme(
        panel.grid = element_blank())+
  geom_vline(data=meanactivenolevel, aes(xintercept=grp.mean, color=Exam), linetype="dashed")

  
  ActivePercentExamnolevel
  #Now to Save!
# - vector based output height = 4.5, width = 6 inches
ggsave(ActivePercentExamnolevel, file=" ActivePercentExamnolevel.png",
       height = 4.5,
       width = 6,
       units = "in",
       dpi = 300)
```

Passive by exam
```{r}
CountPassiveExamnolevel<-CombinedZ %>%
  group_by(Exam) %>%
count(Passive)%>% #This tells me the number of students who said each number of strategies for active by level. 
  rename("number_of_students"="n") %>% #rename the default "value" column as "Percent" 
  mutate(Total_of_students = number_of_students) %>% #I created a dummy column for total number of students. Then, I manually added up all of the lower students and all of the upper students. There are 968 exam 1 students and 845 exam 3 students. 
     mutate(Total_of_students=recode(Total_of_students,"47"="968","330"="968","375"="968", "181"="968", "35"="968", "22"="845","251"="845", "369"="845", "171"="845","32"="845")) %>%
 filter(!is.na(Passive)) %>% #removes any NA for Zexamscore, and removes any exam scores that were 0 before they were z-scored
  mutate(Percent_of_students_Passive = number_of_students) %>%#Create dummy column and fill in values
  mutate(Percent_of_students_Passive=recode(Percent_of_students_Passive, "47"="4.86","330"="34.09","375"="38.74", "181"="18.70", "35"="3.62", "22"="2.60","251"="29.70", "369"="43.67", "171"="20.24","32"="3.79"))%>%
   mutate(Percent_of_students_Passive=as.numeric(Percent_of_students_Passive))%>%
 mutate(Total_of_students=as.numeric(Total_of_students))

meanpassivenolevel<-CombinedZ%>%
  group_by(Exam) %>%
 summarise(mean(Passive, na.rm = TRUE)) %>%
  rename(grp.mean="mean(Passive, na.rm = TRUE)")

PassivePercentExamnolevel<-ggplot(CountPassiveExamnolevel, aes(x=Passive, y=Percent_of_students_Passive, fill=Exam)) +
    geom_col(position = position_dodge(preserve = "single")) +
        scale_fill_manual(values=c("#8491B4B2", "#7E6148B2")) +
   scale_color_manual(values = c("#8491B4B2", "#7E6148B2")) +
    scale_x_continuous(name = "# of study strategies used", breaks=c(0,1,2,3,4,5,6), limits=c(-0.5, 6.5)) + 
    scale_y_continuous(name = "% of students", breaks=c(0, 10, 20, 30, 40), limits=c(0, 45)) +  
    theme_bw() + 
   theme(
        panel.grid = element_blank())+
  geom_vline(data=meanpassivenolevel, aes(xintercept=grp.mean, color=Exam), linetype="dashed")

  
  PassivePercentExamnolevel
  #Now to Save!
# - vector based output height = 4.5, width = 6 inches
ggsave(PassivePercentExamnolevel, file="PassivePercentExamnolevel.png",
       height = 4.5,
       width = 6,
       units = "in",
       dpi = 300)
```

test for significant differences between study habits over time (exam number)

```{r}
#active strats over time count
Activetime<-glmer(Active~Exam + Level + Exam:Level + (1|instructoranon/Student), family=poisson, data=CombinedZ, na.action=na.exclude) 

summary(Activetime)

#No significant interaction, so we can analyze levels together

Activetimenolevel<-glmer(Active~Exam +(1|instructoranon/Student), family=poisson, data=CombinedZ, na.action=na.exclude) 

summary(Activetimenolevel)


#active strats over time count by class
Activetime<-glmer(Active~Exam + Class + Exam:Class + (1|instructoranon/Student), family=poisson, data=CombinedZ, na.action=na.exclude) 

summary(Activetime)

##No significant interaction of class, so removed that. 
Activetime<-glmer(Active~Exam + Class + (1|instructoranon/Student), family=poisson, data=CombinedZ, na.action=na.exclude) 

summary(Activetime)
confint(Activetime)

##Passive by class and time

Passivetime<-glmer(Passive~Exam + Class + Exam:Class + (1|instructoranon/Student), family=poisson, data=CombinedZ, na.action=na.exclude) 

summary(Passivetime)

##No significant interaction of class, so removed that. 
Passivetime<-glmer(Passive~Exam + Class + (1|instructoranon/Student), family=poisson, data=CombinedZ, na.action=na.exclude) 

summary(Passivetime)

confint(Passivetime)




```

